---
title: 'ESM 204: Homework 2'
author: "Anthony Luna, Kazia Mermel, & Keene Morrow"
date: "4/27/2020"
output:
  html_document:
subtitle: Simulating a Carbon Market
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(stringr)
library(boot)
library(broom)

carbon <- read_csv("ESM204_HW2_Data.csv")

carbon_long <- carbon %>%
  pivot_longer(cols = 2:5, names_to = "sector") %>%
  mutate(sector = str_sub(sector, 4)) %>%
  rename(abatement = Abatement,
         mc = value)

ggplot(data = carbon_long) +
  geom_point(aes(x = abatement,
                 y = mc,
                 color = sector)) +
  facet_wrap(~ sector)

# current emissions
current_a <- 170
current_b <- 200
current_c <- 230
current_d <- 300
```

***

To comply with the Paris Accord, governments around the world are seeking ways to reduce carbon emissions. You have been asked to provide guidance on how country “X” can most efficiently reduce carbon emissions. In this homework you will conduct data analysis, simulations, and policy analysis to help inform this decision. The data for this assignment are available in the file “HW2Data.csv”, available on Gauchospace. The data contain estimates of the marginal cost of carbon abatement for four economic sectors (sectors A, B, C, and D), for different levels of abatement. A few notes:

- Current carbon emissions in the sectors are: (A=170, B=200, C=230, and D=300).
- Sectors A, B, and C are in country X. Sector D is in a different country, Y.
- Prices are in $/ton and quantities are in tons.

Your task is to produce a *1 page memo*, with supporting appendices, to advise country X (and in some questions, country Y) about how best to reduce carbon emissions. Your memo and appendices must address the following questions.
1. Start with sector A. Plot that sector’s marginal abatement cost estimates and write down a plausible functional form for the marginal cost of abatement. Use regression analysis to estimate the parameters of that function. Repeating this for sectors B, C, and D will give you a model of the marginal cost of abatement function for each sector. How well do your models fit the data for each sector? You may need to experiment with different functional forms, and you don’t necessarily have to assume the same function for each sector. Produce a plot of the estimated marginal abatement cost functions in all four sectors (this plot should go in your memo).

### 1.

Various models were explored for each sector.  The model with the best fit was selected by comparing R^2 and AIC values for each model, as appropriate.

#### Sector A

```{r include = FALSE}
# subset for just Sector A
carbon_a <- carbon_long %>%
  filter(sector == "A")

# Model Exploration
## Linear
carbon_a_lm <- lm(mc ~ abatement, data = carbon_a)
summary(carbon_a_lm)
AIC(carbon_a_lm)

carbon_a_lm$coefficients[1]
carbon_a_lm$coefficients[2]

# Create Data Frame
predict_carbon_a <- data.frame(mc_predict = predict(carbon_a_lm,carbon_a), sector = "A",
                                   abatement = carbon_a$abatement)

ggplot()+
  geom_point(data = carbon_a,
             aes(x = abatement,
                                 y = mc)) +
  geom_path(data = predict_carbon_a,
            aes(x = abatement,
                y = mc_predict)) +
  labs(y = "Cost ($/ton)",
         x = "Carbon (tons)",
         title = "Sector A",
         # subtitle = "",
         caption = "ESM 204 Spring 2020\nAnthony Luna, Kazia Mermel, & Keene Morrow") + 
    theme_minimal() +
    theme(plot.caption = element_text(hjust = 0, face = "italic"))

```

*Linear model equation:*

mc = `r round(carbon_a_lm$coefficients[2], 2)` * abatement + `r round(carbon_a_lm$coefficients[1], 2)`


#### Sector B
```{r include = FALSE}
# subset for just Sector B
carbon_b <- carbon_long %>%
  filter(sector == "B")

# Model Exploration
## Linear
carbon_b_lm <- lm(mc ~ abatement, data = carbon_b)
carbon_b_lm

## Non-Linear

# estimate growth rate (r) to be = carbon_b_lm$coefficients[2] ~ 0.39
# estimate K to be 60
# A = (K - P) / P = 10

carbon_b_nls <- nls(mc ~ K/(1 + A*exp(-r*abatement)),
             data = carbon_b,
             start = list(K = 58,
                          A = 10,
                          r = carbon_b_lm$coefficients[2]),
             trace = TRUE)

summary(carbon_b_nls)

model_out_b <- broom::tidy(carbon_b_nls) # cleaner model output with broom
model_out_b

## Predicted values using the model for plotting
predict_carbon_b_nls <- data.frame(mc_predict = predict(carbon_b_nls, carbon_b$abatement),
                                   sector = "B",
                                   abatement = carbon_b$abatement)

ggplot()+
  geom_point(data = carbon_b,
             aes(x = abatement,
                                 y = mc)) +
  geom_path(data = predict_carbon_b_nls,
            aes(x = abatement,
                y = mc_predict)) +
  labs(y = "Cost ($/ton)",
         x = "Carbon (tons)",
         title = "Sector B",
         # subtitle = "",
         caption = "ESM 204 Spring 2020\nAnthony Luna, Kazia Mermel, & Keene Morrow") + 
    theme_minimal() +
    theme(plot.caption = element_text(hjust = 0, face = "italic"))
 


```

*Logistic model equation:*

mc = `r round(model_out_b$estimate[1], 2)` /(1 + `r round(model_out_b$estimate[2], 2)` * exp(- `r round(model_out_b$estimate[3], 2)` * abatement)


#### Sector C
```{r include = FALSE}
# subset for just Sector C
carbon_c <- carbon_long %>%
  filter(sector == "C")

# Exploration of Sector C's values abatement > 40
# subset for > 40
carbon_c_40 <- carbon_long %>%
  filter(sector == "C",
         abatement > 39)
# Model Exploration
## Linear - Abatement greater than 40
carbon_c_40_lm <- lm(mc ~ abatement,
                  data = carbon_c_40)

summary(carbon_c_40_lm)
# Multiple R-squared:  0.9149,	Adjusted R-squared:  0.9063 
AIC(carbon_c_40_lm) # 104.4639


# > 40 (blue)

new_df <- data.frame(abatement = carbon_c_40$abatement)

mc_predict <- predict(carbon_c_40_lm, newdata = new_df)

predict_carbon_c_40_lm <- data.frame(new_df, mc_predict) %>% 
  mutate(sector = "C")

predict_c_stage <- data.frame(abatement = seq(0,30,by=10),sector = rep("C",4), mc_predict = rep(0,4))

predict_carbon_c_40_lm <- predict_carbon_c_40_lm %>% 
  union(predict_c_stage) %>% 
  arrange(abatement)


ggplot()+
  geom_point(data = carbon_c,
             aes(x = abatement,
                 y = mc)) +
  geom_path(data = predict_carbon_c_40_lm,
            aes(x = abatement,
                y = mc_predict),
            color = "blue") +
  labs(y = "Cost ($/ton)",
         x = "Carbon (tons)",
         title = "Sector C",
         # subtitle = "",
         caption = "ESM 204 Spring 2020\nAnthony Luna, Kazia Mermel, & Keene Morrow") + 
    theme_minimal() +
    theme(plot.caption = element_text(hjust = 0, face = "italic"))
```

*Piecewise linear model equation:*

From 0 to `r -round((carbon_c_40_lm$coefficients[1] / carbon_c_40_lm$coefficients[2]), 2)`: mc = 0

From `r -round((carbon_c_40_lm$coefficients[1] / carbon_c_40_lm$coefficients[2]), 2)` to infinity: mc = `r round(carbon_c_40_lm$coefficients[2], 2)` * abatement - `r -round(carbon_c_40_lm$coefficients[1], 2)`

#### Sector D
```{r include = FALSE}
# subset for just Sector C
carbon_d <- carbon_long %>%
  filter(sector == "D")

# Model Exploration
## Non-Linear

# estimate growth rate (r) to be = carbon_d_lm$coefficients[2] ~ 
# estimate K to be 85
# A = (K - P) / P = 
carbon_d_lm <- lm(mc ~abatement,data=carbon_d)
carbon_d_nls <- nls(mc ~ K/(1 + A*exp(-r*abatement)),
                    data = carbon_d,
                    start = list(K = 112,
                                 A = 68,
                                 r = carbon_d_lm$coefficients[2]),
                    trace = TRUE)

summary(carbon_d_nls) # Residual standard error: 9.089 on 13 degrees of freedom
AIC(carbon_d_nls) # 120.7116

model_out_d <- broom::tidy(carbon_d_nls) # cleaner model output with broom
model_out_d

## Predicted values using the nls model for plotting
predict_carbon_d_nls <- data.frame(mc_predict = predict(carbon_d_nls, carbon_d$abatement),
                                   sector = "D",
                                   abatement = carbon_d$abatement)


ggplot()+
  geom_point(data = carbon_d,
             aes(x = abatement,
                 y = mc)) +
  geom_path(data = predict_carbon_d_nls,
            aes(x = abatement,
                y = mc_predict)) +
  labs(y = "Cost ($/ton)",
         x = "Carbon (tons)",
         title = "Sector D",
         # subtitle = "",
         caption = "ESM 204 Spring 2020\nAnthony Luna, Kazia Mermel, & Keene Morrow") + 
    theme_minimal() +
    theme(plot.caption = element_text(hjust = 0, face = "italic"))


```

*Logistic model equation:*

mc = `r round(model_out_d$estimate[1], 2)`/(1 + `r round(model_out_d$estimate[2], 2)` * exp(- `r round(model_out_d$estimate[3], 2)` * abatement)


#### Demand Curves

The above functions were used to calculate values for demand in a range of possible levels of abatement/pollution from 0 to 300 tons of carbon.  These values were then plotted in Figure 1.
```{r}
carbon_a_fun <- function(Q){
  m <- as.numeric(carbon_a_lm$coefficients[2])
  b <- as.numeric(carbon_a_lm$coefficients[1])
  m*Q+b}

carbon_a_fun_demand <- function(Q) {carbon_a_fun(current_a-Q)}

carbon_b_fun <- function(Q){
  K <- as.numeric(coefficients(carbon_b_nls)[1])
  A <- as.numeric(coefficients(carbon_b_nls)[2])
  r <- as.numeric(coefficients(carbon_b_nls)[3])
  K/(1 + A*exp(-r*Q))
     }

carbon_b_fun_demand <- function(Q) {carbon_b_fun(current_b-Q)}

carbon_c_fun <- function(Q){
  m <- as.numeric(carbon_c_40_lm$coefficients[2])
  b <- as.numeric(carbon_c_40_lm$coefficients[1])
  ifelse(Q<40,0,m*Q+b)}

carbon_c_fun_demand <- function(Q){ifelse(Q<40,carbon_c_fun(current_c),carbon_c_fun(current_c-Q+40))}

carbon_d_fun <- function(Q){
  K <- as.numeric(coefficients(carbon_d_nls)[1])
  A <- as.numeric(coefficients(carbon_d_nls)[2])
  r <- as.numeric(coefficients(carbon_d_nls)[3])
  K/(1 + A*exp(-r*Q))
     }

carbon_d_fun_demand <- function(Q) {carbon_d_fun(current_d-Q)}
```


```{r}
abate_range <- seq(0, 300, by = 1)

predict_supply <-  data.frame(abatement= abate_range, 
  type = rep("Supply",
             length(abate_range)),
  sector_a = carbon_a_fun(abate_range),
  sector_b = carbon_b_fun(abate_range),
  sector_c = carbon_c_fun(abate_range),
  sector_d = carbon_d_fun(abate_range)) %>% 
  pivot_longer(cols = 3:6,
               values_to = "mc_predict",
               names_to = "sector",
               names_prefix = "sector_") 

predict_demand <- data.frame(
  abatement= abate_range,
  type = rep("Demand",
             length(abate_range)),
  sector_a= carbon_a_fun_demand(abate_range),
  sector_b= carbon_b_fun_demand(abate_range),
  sector_c= carbon_c_fun_demand(abate_range),
  sector_d= carbon_d_fun_demand(abate_range) ) %>% 
  pivot_longer(cols = 3:6,
               values_to = "mc_predict",
               names_to = "sector",
               names_prefix = "sector_")

predict_values <- predict_supply %>% 
  union(predict_demand) %>% 
  arrange(sector,
          type,
          abatement)

predict_plot <- predict_values %>%
  mutate(sector = case_when(
    sector == "a" ~ "A",
    sector == "b" ~ "B",
    sector == "c" ~ "C",
    sector == "d" ~ "D"
  )) %>%
  filter(mc_predict >= 0)
```

### 2. 

#### Supply Curves

The above functions were used to calculate values for supply in a range of possible levels of abatement/pollution from 0 to 300 tons of carbon.  These values were then plotted in Figure 1.

```{r supply_demand_plot}
ggplot(data = predict_plot) +
  geom_path(aes(x = abatement,
                y = mc_predict,
                color = type)) +
 facet_wrap(~sector,
             scales = "free")  +
  labs(y = "Cost ($/ton)",
       x = "Carbon (tons)",
       title = "Supply & Demand by Sector",
       # subtitle = "",
       caption = "Figure 1.\nESM 204 Spring 2020\nAnthony Luna, Kazia Mermel, & Keene Morrow") + 
  theme_minimal() +
  theme(plot.caption = element_text(hjust = 0, face = "italic"),
        legend.title = element_blank())
  
```

### 3.

#### a.

```{r}
# (1) Total Cost
total_cost <- integrate(carbon_a_fun,0,100)$value + integrate(carbon_b_fun,0,100)$value + integrate(carbon_c_fun,0,100)$value
# (2) Cost or Benefit to each sector (negative is cost)
sector_a_cost <-integrate(carbon_a_fun,0,100)$value
sector_b_cost <-integrate(carbon_b_fun,0,100)$value
sector_c_cost <-integrate(carbon_c_fun,0,100)$value
# (3) There won't be any tax revenue here...
```

The cost of abatement for each sector (A, B, and C) was calculated by integrating the functions derived previously from 0 to 100 tons of carbon.  Total cost was calculated as the sum of all sectors.

#### b.

```{r}
# Inverse all three demand functions and handle edge cases (changing NAs to Zero etc)
carbon_a_fun_inv <- function(P){
  m <- as.numeric(carbon_a_lm$coefficients[2])
  b <- as.numeric(carbon_a_lm$coefficients[1])
  current_a-((P-b)/m)}


carbon_b_fun_inv <- function(P){
  K <- as.numeric(coefficients(carbon_b_nls)[1])
  A <- as.numeric(coefficients(carbon_b_nls)[2])
  r <- as.numeric(coefficients(carbon_b_nls)[3])
  ifelse(is.na(current_b-(-(1/r)*log((1/A)*((K/P)-1)))),0,current_b-(-(1/r)*log((1/A)*((K/P)-1))))
     }


carbon_c_fun_inv <- function(P){
  m <- as.numeric(carbon_c_40_lm$coefficients[2])
  b <- as.numeric(carbon_c_40_lm$coefficients[1])
  ifelse(P<=carbon_c_fun_demand(0),current_c-((P-b)/m)+40,0)
}


carbon_d_fun_inv <- function(P){
  K <- as.numeric(coefficients(carbon_d_nls)[1])
  A <- as.numeric(coefficients(carbon_d_nls)[2])
  r <- as.numeric(coefficients(carbon_d_nls)[3])
  current_d-(-(1/r)*log((1/A)*((K/P)-1)))
     }
```


```{r}
#Construct dataframe with dummy data between relevant P range where the final column give the aggregate
mc_range <- seq(0,350,by=.01)
carbon_agg <- data.frame(mc= mc_range,
  sector_a = ifelse(carbon_a_fun_inv(mc_range)>0,carbon_a_fun_inv(mc_range),0),
  sector_b = ifelse(carbon_b_fun_inv(mc_range)>0,carbon_b_fun_inv(mc_range),0),
  sector_c = carbon_c_fun_inv(mc_range)) %>% 
  mutate(agg = sector_a+sector_b+sector_c)
```


```{r}
# find tax by looking at when agg = 300

tax <- carbon_agg %>% filter(round(agg)==300) %>% 
  summarize(mc=mean(mc),
            sector_a=mean(sector_a),
            sector_b=mean(sector_b),
            sector_c=mean(sector_c),
            agg=mean(agg)
            )


# Plot to check that the aggregate values look about right...
# ggplot(data = predict_values %>%
#          filter(type == "Demand") %>%
#          filter(sector!= "d")) +
#   geom_path(aes(x = abatement,
#                 y = mc_predict,
#                 color = sector)) +
#   ylim(0,300) +
#   xlim(0,600)

ggplot() +
  geom_path(data = carbon_agg, 
            aes(x = agg,
                y = mc))+
  geom_path(data = carbon_agg %>%
              mutate(agg_sup = agg),
            aes(x = agg,
                y = mc))+
  # Here are some paths of the sectors if that would be helpful for a graph...
  geom_path(data = carbon_agg, aes(x = sector_a, y = mc), color = "red") +
  geom_path(data = carbon_agg, aes(x = sector_b, y = mc), color = "dark green") +
  geom_path(data = carbon_agg, aes(x = sector_c, y = mc), color = "blue") +
  # here is the tax line...
  geom_hline(yintercept = tax$mc,
             linetype = "dashed",
             color = "gray")+
  xlim(0,600)   +
  labs(y = "Cost ($/ton)",
       x = "Carbon (tons)",
       title = "Emission Reduction Through Taxation",
       subtitle = "Demand & Tax Level",
       caption = "ESM 204 Spring 2020\nAnthony Luna, Kazia Mermel, & Keene Morrow") + 
  theme_minimal() +
  theme(plot.caption = element_text(hjust = 0, face = "italic")) +
  # Labels
  geom_text(aes(x = 50,
                y = 25),
            label = "Sector A",
            color = "red",
            hjust = 0) +
  geom_text(aes(x = 100,
                y = 75),
            label = "Sector B",
            color = "dark green",
            hjust = 0) +
  geom_text(aes(x = 100,
                y = 200),
            label = "Sector C",
            color = "blue",
            hjust = 0) +
  geom_text(aes(x = 500,
                y = 35),
            label = "Aggregate",
            color = "black",
            hjust = 0) +
  geom_text(aes(x = 500,
                y = 75),
            label = "Tax",
            color = "grey",
            hjust = 0)

```


```{r tax_answer}
#(1) Total cost 
# The way this works is that the tax rate that we found
# is used to calculate the demand for a given sector. 
# Given this demand, we know that the benefit lost (or cost)
# is equal to the area under the demand curve for a given 
# sector going from the demand given the tax to the amount 
# produced before the tax.

total_cost_tax <-  integrate(carbon_a_fun_demand, carbon_a_fun_inv(tax$mc),current_a)$value+
  integrate(carbon_b_fun_demand,carbon_b_fun_inv(tax$mc),current_b)$value+
  integrate(carbon_c_fun_demand,carbon_c_fun_inv(tax$mc),current_c)$value

#(2) Sector Cost This is just the total cost broken up by sector

sector_a_cost_tax <- integrate(carbon_a_fun_demand,carbon_a_fun_inv(tax$mc),current_a)$value
sector_b_cost_tax <- integrate(carbon_b_fun_demand,carbon_b_fun_inv(tax$mc),current_b)$value
sector_c_cost_tax <- integrate(carbon_c_fun_demand,carbon_c_fun_inv(tax$mc),current_c)$value

#(3) There are a couple ways to calculate this but this is the easiest.
tax_rev <- tax$agg*tax$mc

```

The tax to produce the desired reduction was calculated by expressing the functions for the sectors of interest in terms of abatement/pollution, then calculating the price where for the appropriate abatement level (300 tons). This allowed the total cost of the tax to be calculated by integrating each sector's function from the desired carbon emission that would be accomplished by the tax to the quantity of carbon currently emitted.  The total cost of the tax was found by summing those values and the tax revenue by multiplying the tax (\$`r round(tax$mc, 2)`/ton) by the emission level (300 tons).


#### c.

Using the tax calculated above as the price to achieve a reduction of 300 tons of carbon emissions, the demand for carbon for each sector was calculated.  The difference between the allocated 100 tons of carbon emissions and the demand was calculated, then used to determine the cost or benefit to each sector at their respective equilibriums  The total cost of the program was calculated and found to be the same as the total cost of the tax in part b.  There is no tax revenue associated with this scenario.

```{r}
#Allocated at start
credit <- 100

#Demand for credits
a_cnt_demand <- carbon_a_fun_inv(tax$mc)
b_cnt_demand <- carbon_b_fun_inv(tax$mc)
c_cnt_demand <- carbon_c_fun_inv(tax$mc)

#Surplus or deficit of credits
a_cnt_sold <- credit - a_cnt_demand
b_cnt_sold <- credit - b_cnt_demand
c_cnt_sold <- credit - c_cnt_demand

# (1) Total Cost
# This is similar to the tax question
total_cost_cnt <-integrate(carbon_a_fun_demand, carbon_a_fun_inv(tax$mc),current_a)$value+
  integrate(carbon_b_fun_demand,carbon_b_fun_inv(tax$mc),current_b)$value+
  integrate(carbon_c_fun_demand,carbon_c_fun_inv(tax$mc),current_c)$value

# (2) Sector cost 
# This is similar to the tax question. Add the extra area to calculation such that the sold/purchased credits are accounted for
sector_a_cost_cnt <-  integrate(carbon_a_fun_demand,carbon_a_fun_inv(tax$mc),current_a)$value-
  integrate(carbon_a_fun_demand,carbon_a_fun_inv(tax$mc),100)$value
sector_b_cost_cnt <-  integrate(carbon_b_fun_demand,carbon_b_fun_inv(tax$mc),current_b)$value-
  integrate(carbon_b_fun_demand,carbon_b_fun_inv(tax$mc),100)$value
sector_c_cost_cnt <-  integrate(carbon_c_fun_demand,carbon_c_fun_inv(tax$mc),current_c)$value+
  integrate(carbon_a_fun_demand,carbon_a_fun_inv(tax$mc),100)$value+
  integrate(carbon_b_fun_demand,carbon_b_fun_inv(tax$mc),100)$value

# (3) No tax revenue
```

### 4.

```{r}
# Carbon_agg as calculated above, but with sector d included
carbon_agg_d <- data.frame(mc= mc_range,
  sector_a = ifelse(carbon_a_fun_inv(mc_range)>0,carbon_a_fun_inv(mc_range),0),
  sector_b = ifelse(carbon_b_fun_inv(mc_range)>0,carbon_b_fun_inv(mc_range),0),
  sector_c = carbon_c_fun_inv(mc_range),
  sector_d = ifelse(carbon_d_fun_inv(mc_range)>0,carbon_d_fun_inv(mc_range),0)) %>% 
  mutate(agg = sector_a+sector_b+sector_c+sector_d)


# named tax_d for convention but really gives the socailly optimal 
# price of carbon under this scenario. THere is a decrease in the 
# price meaning higher reductions possibly?
tax_d <- carbon_agg_d %>% filter(round(agg)==600) %>% 
  summarize(mc=mean(mc),
            sector_a=mean(sector_a),
            sector_b=mean(sector_b),
            sector_c=mean(sector_c),
            sector_d=mean(sector_d),
            agg=mean(agg)
            )
```

