---
title: "ESM 204: Homework 2"
subtitle: "Simulating a Carbon Market"
author: "Anthony Luna, Kazia Mermel, & Keene Morrow"
date: "4/27/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(stringr)
library(boot)
library(broom)

carbon <- read_csv("ESM204_HW2_Data.csv")

carbon_long <- carbon %>%
  pivot_longer(cols = 2:5, names_to = "sector") %>%
  mutate(sector = str_sub(sector, 4)) %>%
  rename(abatement = Abatement,
         mc = value)

ggplot(data = carbon_long) +
  geom_point(aes(x = abatement,
                 y = mc,
                 color = sector)) +
  facet_wrap(~ sector)

# current emissions
current_a <- 170
current_b <- 200
current_c <- 230
current_d <- 300
```

# Saving to Git
1. Save your shit
2. Go to the Git tab
3. Click the little white boxes next to the stuff you want to send to the other person
4. Click Commit
5. Write a witty but descriptive commit message
6. Hit Commit!
7. *PUUUUUUULLLLLLL* then push (I enjoy how much this sounds like instructions for a dance move)
8. Profit

***

To comply with the Paris Accord, governments around the world are seeking ways to reduce carbon emissions. You have been asked to provide guidance on how country “X” can most efficiently reduce carbon emissions. In this homework you will conduct data analysis, simulations, and policy analysis to help inform this decision. The data for this assignment are available in the file “HW2Data.csv”, available on Gauchospace. The data contain estimates of the marginal cost of carbon abatement for four economic sectors (sectors A, B, C, and D), for different levels of abatement. A few notes:

- Current carbon emissions in the sectors are: (A=170, B=200, C=230, and D=300).
- Sectors A, B, and C are in country X. Sector D is in a different country, Y.
- Prices are in $/ton and quantities are in tons.

Your task is to produce a *1 page memo*, with supporting appendices, to advise country X (and in some questions, country Y) about how best to reduce carbon emissions. Your memo and appendices must address the following questions.
1. Start with sector A. Plot that sector’s marginal abatement cost estimates and write down a plausible functional form for the marginal cost of abatement. Use regression analysis to estimate the parameters of that function. Repeating this for sectors B, C, and D will give you a model of the marginal cost of abatement function for each sector. How well do your models fit the data for each sector? You may need to experiment with different functional forms, and you don’t necessarily have to assume the same function for each sector. Produce a plot of the estimated marginal abatement cost functions in all four sectors (this plot should go in your memo).

# A - Choose first order polynomial (Linear)
```{r}
# subset for just Sector A
carbon_a <- carbon_long %>%
  filter(sector == "A")


# Model Exploration
## Linear
carbon_a_lm <- lm(mc ~ abatement, data = carbon_a)
summary(carbon_a_lm)

carbon_a_lm$coefficients[1]
carbon_a_lm$coefficients[2]

# Create Data Frame
predict_carbon_a <- data.frame(mc_predict = predict(carbon_a_lm,carbon_a), sector = "A",
                                   abatement = carbon_a$abatement)

```

*Sector A Equation is:*
mc = `r carbon_a_lm$coefficients[2]` * abatement + `r carbon_a_lm$coefficients[1]`


# B - Choose Logistic Model
```{r}
# subset for just Sector B
carbon_b <- carbon_long %>%
  filter(sector == "B")

# Model Exploration
## Linear
carbon_b_lm <- lm(mc ~ abatement, data = carbon_b)
carbon_b_lm

## Non-Linear

# estimate growth rate (r) to be = carbon_b_lm$coefficients[2] ~ 0.39
# estimate K to be 60
# A = (K - P) / P = 10

carbon_b_nls <- nls(mc ~ K/(1 + A*exp(-r*abatement)),
             data = carbon_b,
             start = list(K = 58,
                          A = 10,
                          r = carbon_b_lm$coefficients[2]),
             trace = TRUE)

summary(carbon_b_nls)

model_out_b <- broom::tidy(carbon_b_nls) # cleaner model output with broom
model_out_b

## Predicted values using the model for plotting
predict_carbon_b_nls <- data.frame(mc_predict = predict(carbon_b_nls, carbon_b$abatement),
                                   sector = "B",
                                   abatement = carbon_b$abatement)

ggplot()+
  geom_point(data = carbon_b,
             aes(x = abatement,
                                 y = mc)) +
  geom_path(data = predict_carbon_b_nls,
            aes(x = abatement,
                y = mc_predict))
 


```

*Sector B Equation is:*
mc = K/(1 + A * exp(- r * abatement)

where

K = `r model_out_b$estimate[1]`

A = `r model_out_b$estimate[2]`

r = `r model_out_b$estimate[3]`

# C - Choose Piecewise
```{r}
# subset for just Sector C
carbon_c <- carbon_long %>%
  filter(sector == "C")

# Exploration of Sector C's values abatement > 40
# subset for > 40
carbon_c_40 <- carbon_long %>%
  filter(sector == "C",
         abatement > 39)
# Model Exploration
## Linear - Abatement greater than 40
carbon_c_40_lm <- lm(mc ~ abatement,
                  data = carbon_c_40)

summary(carbon_c_40_lm)
# Multiple R-squared:  0.9149,	Adjusted R-squared:  0.9063 
AIC(carbon_c_40_lm) # 104.4639


# > 40 (blue)

new_df <- data.frame(abatement = carbon_c_40$abatement)

mc_predict <- predict(carbon_c_40_lm, newdata = new_df)

predict_carbon_c_40_lm <- data.frame(new_df, mc_predict) %>% 
  mutate(sector = "C")

predict_c_stage <- data.frame(abatement = seq(0,30,by=10),sector = rep("C",4), mc_predict = rep(0,4))

predict_carbon_c_40_lm <- predict_carbon_c_40_lm %>% 
  union(predict_c_stage) %>% 
  arrange(abatement)

# Comparison
ggplot()+
  geom_point(data = carbon_c,
             aes(x = abatement,
                 y = mc)) +
  geom_path(data = predict_carbon_c_40_lm,
            aes(x = abatement,
                y = mc_predict),
            color = "blue")

# lm for values > 40 has lowest AIC (104 vs 137 for lowest full data set model)

```

*Sector C Equation is:*

From 0 to `r -carbon_c_40_lm$coefficients[1] / carbon_c_40_lm$coefficients[2]`:

mc = 0

From `r -carbon_c_40_lm$coefficients[1] / carbon_c_40_lm$coefficients[2]` to infinity:

mc = `r carbon_c_40_lm$coefficients[2]` * abatement + `r carbon_c_40_lm$coefficients[1]`

# Sector D - Choose Logistic Model
```{r}
# subset for just Sector C
carbon_d <- carbon_long %>%
  filter(sector == "D")

# Model Exploration
## Non-Linear

# estimate growth rate (r) to be = carbon_d_lm$coefficients[2] ~ 
# estimate K to be 85
# A = (K - P) / P = 
carbon_d_lm <- lm(mc ~abatement,data=carbon_d)
carbon_d_nls <- nls(mc ~ K/(1 + A*exp(-r*abatement)),
                    data = carbon_d,
                    start = list(K = 112,
                                 A = 68,
                                 r = carbon_d_lm$coefficients[2]),
                    trace = TRUE)

summary(carbon_d_nls) # Residual standard error: 9.089 on 13 degrees of freedom
AIC(carbon_d_nls) # 120.7116

model_out_d <- broom::tidy(carbon_d_nls) # cleaner model output with broom
model_out_d

## Predicted values using the nls model for plotting
predict_carbon_d_nls <- data.frame(mc_predict = predict(carbon_d_nls, carbon_d$abatement),
                                   sector = "D",
                                   abatement = carbon_d$abatement)


ggplot()+
  geom_point(data = carbon_d,
             aes(x = abatement,
                 y = mc)) +
  geom_path(data = predict_carbon_d_nls,
            aes(x = abatement,
                y = mc_predict)) +
  labs(y = "Cost ($/ton)",
         x = "Abatement (tons)",
         title = "Sector D",
         # subtitle = "",
         caption = "ESM 204 Spring 2020\nAnthony Luna, Kazia Mermel, & Keene Morrow") + 
    theme_minimal() +
    theme(plot.caption = element_text(hjust = 0, face = "italic"))


```

*Sector D Equation is:*
mc = K/(1 + A * exp(- r * abatement)

where

K = `r model_out_d$estimate[1]`

A = `r model_out_d$estimate[2]`

r = `r model_out_d$estimate[3]`




```{r}
carbon_a_fun <- function(Q){
  m <- as.numeric(carbon_a_lm$coefficients[2])
  b <- as.numeric(carbon_a_lm$coefficients[1])
  m*Q+b}

carbon_a_fun_demand <- function(Q) {carbon_a_fun(current_a-Q)}

carbon_b_fun <- function(Q){
  K <- as.numeric(coefficients(carbon_b_nls)[1])
  A <- as.numeric(coefficients(carbon_b_nls)[2])
  r <- as.numeric(coefficients(carbon_b_nls)[3])
  K/(1 + A*exp(-r*Q))
     }

carbon_b_fun_demand <- function(Q) {carbon_b_fun(current_b-Q)}

carbon_c_fun <- function(Q){
  m <- as.numeric(carbon_c_40_lm$coefficients[2])
  b <- as.numeric(carbon_c_40_lm$coefficients[1])
  ifelse(Q<40,0,m*Q+b)}

carbon_c_fun_demand <- function(Q){ifelse(Q<40,carbon_c_fun(current_c),carbon_c_fun(current_c-Q+40))}

carbon_d_fun <- function(Q){
  K <- as.numeric(coefficients(carbon_d_nls)[1])
  A <- as.numeric(coefficients(carbon_d_nls)[2])
  r <- as.numeric(coefficients(carbon_d_nls)[3])
  K/(1 + A*exp(-r*Q))
     }

carbon_d_fun_demand <- function(Q) {carbon_d_fun(current_d-Q)}
```


```{r}
abate_range <- seq(0, 300, by = 1)

predict_supply <-  data.frame(abatement= abate_range, 
  type = rep("Supply",
             length(abate_range)),
  sector_a = carbon_a_fun(abate_range),
  sector_b = carbon_b_fun(abate_range),
  sector_c = carbon_c_fun(abate_range),
  sector_d = carbon_d_fun(abate_range)) %>% 
  pivot_longer(cols = 3:6,
               values_to = "mc_predict",
               names_to = "sector",
               names_prefix = "sector_") 

predict_demand <- data.frame(
  abatement= abate_range,
  type = rep("Demand",
             length(abate_range)),
  sector_a= carbon_a_fun_demand(abate_range),
  sector_b= carbon_b_fun_demand(abate_range),
  sector_c= carbon_c_fun_demand(abate_range),
  sector_d= carbon_d_fun_demand(abate_range) ) %>% 
  pivot_longer(cols = 3:6,
               values_to = "mc_predict",
               names_to = "sector",
               names_prefix = "sector_")

predict_values <- predict_supply %>% 
  union(predict_demand) %>% 
  arrange(sector,
          type,
          abatement)

predict_plot <- predict_values %>%
  mutate(sector = case_when(
    sector == "a" ~ "A",
    sector == "b" ~ "B",
    sector == "c" ~ "C",
    sector == "d" ~ "D"
  )) %>%
  filter(mc_predict >= 0)
```


```{r supply_demand_plot}
ggplot(data = predict_plot) +
  geom_path(aes(x = abatement,
                y = mc_predict,
                color = type)) +
 facet_wrap(~sector,
             scales = "free")  +
  labs(y = "Cost ($/ton)",
       x = "Carbon (tons)",
       title = "Supply & Demand by Sector",
       # subtitle = "",
       caption = "Figure 1.\nESM 204 Spring 2020\nAnthony Luna, Kazia Mermel, & Keene Morrow") + 
  theme_minimal() +
  theme(plot.caption = element_text(hjust = 0, face = "italic"),
        legend.title = element_blank())
  
```



2. Using these models (i.e. use the functions you have estimated, not the data themselves) and the current level of carbon emissions (provided above), derive each sector’s demand curve for carbon emissions. In other words, how much would each sector be willing to pay for the right to pollute the first ton, second ton, etc.? Plot each sector’s demand curve for carbon emissions on a graph.


Here, we converted from abatement cost to pollution demand. To do this, we substituted (Q0 - Q) for the abatement term.

# A

MCa = 0.513A + 1.662
MWTPa = 0.513*(Q0-Q) + 1.662
MWTPa = 0.513*(170 - Q) + 1.662


```{r}
# Convert from abatement cost to pollution demand
# Substitute (Q0-Q) for abatement

# this chunk is getting grouchy with me so I just wrote out the equations above rather than trying to make them work as functions. If anyone knows how to do that, go for it! -Kazia





```



3. Now focus on country X (which contains sectors A, B, and C). Assume there are no “co-benefits” from carbon abatement (i.e. that there are no local pollutants that are affected by carbon abatement). Suppose to meet the Paris Accord commitments, country X needs to cut all carbon emissions in half. For each of the policy options listed below, derive: (1) the total cost of meeting the target in country X, (2) the cost (or benefit) to each sector, and (3) the tax revenue generated.

a. Cap on carbon. Each sector (A, B, and C) must cut its carbon emissions by 100 tons (thus reducing total emissions from 600 down to 300).

```{r}
# (1) Total Cost
total_cost <- integrate(carbon_a_fun,0,100)$value + integrate(carbon_b_fun,0,100)$value + integrate(carbon_c_fun,0,100)$value
# (2) Cost or Benefit to each sector (negative is cost)
sector_a_cost <-integrate(carbon_a_fun,0,100)$value
sector_b_cost <-integrate(carbon_b_fun,0,100)$value
sector_c_cost <-integrate(carbon_c_fun,0,100)$value
# (3) There won't be any tax revenue here...
```



b. Tax on carbon. To emit a ton of carbon in country X, you must pay a tax of $t. You will need to find the tax that accomplishes the desired reduction.

*Notes from Section*
**Aggregate (Qa+Qb+Qc) must equal emissions goal. Find tax based off that value** 
**Linear model is okay as long as you say it's not the least R^2 value. Sam may give bonus points for non-linear**
```{r}
# Inverse all three demand functions and handle edge cases (changing NAs to Zero etc)
carbon_a_fun_inv <- function(P){
  m <- as.numeric(carbon_a_lm$coefficients[2])
  b <- as.numeric(carbon_a_lm$coefficients[1])
  current_a-((P-b)/m)}


carbon_b_fun_inv <- function(P){
  K <- as.numeric(coefficients(carbon_b_nls)[1])
  A <- as.numeric(coefficients(carbon_b_nls)[2])
  r <- as.numeric(coefficients(carbon_b_nls)[3])
  ifelse(is.na(current_b-(-(1/r)*log((1/A)*((K/P)-1)))),0,current_b-(-(1/r)*log((1/A)*((K/P)-1))))
     }


carbon_c_fun_inv <- function(P){
  m <- as.numeric(carbon_c_40_lm$coefficients[2])
  b <- as.numeric(carbon_c_40_lm$coefficients[1])
  ifelse(P<=carbon_c_fun_demand(0),current_c-((P-b)/m)+40,0)
}


carbon_d_fun_inv <- function(P){
  K <- as.numeric(coefficients(carbon_d_nls)[1])
  A <- as.numeric(coefficients(carbon_d_nls)[2])
  r <- as.numeric(coefficients(carbon_d_nls)[3])
  current_d-(-(1/r)*log((1/A)*((K/P)-1)))
     }
```


```{r}
#Construct dataframe with dummy data between relevant P range where the final column give the aggregate
mc_range <- seq(0,350,by=.01)
carbon_agg <- data.frame(mc= mc_range,
  sector_a = ifelse(carbon_a_fun_inv(mc_range)>0,carbon_a_fun_inv(mc_range),0),
  sector_b = ifelse(carbon_b_fun_inv(mc_range)>0,carbon_b_fun_inv(mc_range),0),
  sector_c = carbon_c_fun_inv(mc_range)) %>% 
  mutate(agg = sector_a+sector_b+sector_c)
```


```{r}
# find tax by looking at when agg = 300

tax <- carbon_agg %>% filter(round(agg)==300) %>% 
  summarize(mc=mean(mc),
            sector_a=mean(sector_a),
            sector_b=mean(sector_b),
            sector_c=mean(sector_c),
            agg=mean(agg)
            )

# Plot to check that the aggregate values look about right...
ggplot(data = predict_values %>% filter(type=="Demand") %>% filter(sector!="d"))+
  geom_path(aes(x=abatement,y=mc_predict,color=sector))+
  ylim(0,300)+xlim(0,600)

ggplot() +
  geom_path(data=carbon_agg, aes(x=agg,y=mc))+
  geom_path(data=carbon_agg %>% mutate(agg_sup=agg), aes(x=agg,y=mc))+
  # Here are some paths of the sectors if that would be helpful for a graph...
  geom_path(data=carbon_agg, aes(x=sector_a,y=mc),color="red") +
  geom_path(data=carbon_agg, aes(x=sector_b,y=mc),color="dark green") +
  geom_path(data=carbon_agg, aes(x=sector_c,y=mc),color="blue") +
  # here is the tax line...
  geom_hline(yintercept = tax$mc,linetype="dashed",color="gray")+
  xlim(0,600)   +
  labs(y = "Cost ($/ton)",
       x = "Carbon (tons)",
       title = "",
       # subtitle = "",
       caption = "ESM 204 Spring 2020\nAnthony Luna, Kazia Mermel, & Keene Morrow") + 
  theme_minimal() +
  theme(plot.caption = element_text(hjust = 0, face = "italic")) +
  # Labels
  geom_text(aes(x = 50,
                y = 25),
            label = "Sector A",
            color = "red",
            hjust = 0) +
  geom_text(aes(x = 100,
                y = 75),
            label = "Sector B",
            color = "dark green",
            hjust = 0) +
  geom_text(aes(x = 100,
                y = 200),
            label = "Sector C",
            color = "blue",
            hjust = 0) +
  geom_text(aes(x = 500,
                y = 35),
            label = "Aggregate",
            color = "black",
            hjust = 0) +
  geom_text(aes(x = 500,
                y = 75),
            label = "Tax",
            color = "grey",
            hjust = 0)



```


```{r tax_answer}
#(1) Total cost 
# The way this works is that the tax rate that we found
# is used to calculate the demand for a given sector. 
# Given this demand, we know that the benefit lost (or cost)
# is equal to the area under the demand curve for a given 
# sector going from the demand given the tax to the amount 
# produced before the tax.

total_cost_tax <-  integrate(carbon_a_fun_demand, carbon_a_fun_inv(tax$mc),current_a)$value+
  integrate(carbon_b_fun_demand,carbon_b_fun_inv(tax$mc),current_b)$value+
  integrate(carbon_c_fun_demand,carbon_c_fun_inv(tax$mc),current_c)$value

#(2) Sector Cost This is just the total cost broken up by sector

sector_a_cost_tax <- integrate(carbon_a_fun_demand,carbon_a_fun_inv(tax$mc),current_a)$value
sector_b_cost_tax <- integrate(carbon_b_fun_demand,carbon_b_fun_inv(tax$mc),current_b)$value
sector_c_cost_tax <- integrate(carbon_c_fun_demand,carbon_c_fun_inv(tax$mc),current_c)$value

#(3) There are a couple ways to calculate this but this is the easiest.
tax_rev <- tax$agg*tax$mc

```

c. Cap and trade. Each sector (A, B, C) is allocated carbon permits equal to their current emissions minus 100 (same as in (a) above), thus achieving the total reduction of 300. Then, these three sectors are allowed to trade with each other. You will need to derive the outcome of that trading.

The calculated tax above is the idealized price to achieve reduction of 300. Therefore we use this as the universal price of carbon to see emissions of each sector. Based on this, there will be a surplus of carbon credits for some sectors, and a deficit for other.
```{r}
#Allocated at start
credit <- 100

#Demand for credits
a_cnt_demand <- carbon_a_fun_inv(tax$mc)
b_cnt_demand <- carbon_b_fun_inv(tax$mc)
c_cnt_demand <- carbon_c_fun_inv(tax$mc)

#Surplus or deficit of credits
a_cnt_sold <- credit - a_cnt_demand
b_cnt_sold <- credit - b_cnt_demand
c_cnt_sold <- credit - c_cnt_demand

# (1) Total Cost
# This is similar to the tax question
total_cost_cnt <-integrate(carbon_a_fun_demand, carbon_a_fun_inv(tax$mc),current_a)$value+
  integrate(carbon_b_fun_demand,carbon_b_fun_inv(tax$mc),current_b)$value+
  integrate(carbon_c_fun_demand,carbon_c_fun_inv(tax$mc),current_c)$value

# (2) Sector cost
# This is similar to the tax question. Add the extra area to calculation such that the sold/purchased credits are accounted for
sector_a_cost_cnt <-  integrate(carbon_a_fun_demand,carbon_a_fun_inv(tax$mc),current_a)$value-
  integrate(carbon_a_fun_demand,carbon_a_fun_inv(tax$mc),100)$value
sector_b_cost_cnt <-  integrate(carbon_b_fun_demand,carbon_b_fun_inv(tax$mc),current_b)$value-
  integrate(carbon_b_fun_demand,carbon_b_fun_inv(tax$mc),100)$value
sector_c_cost_cnt <-  integrate(carbon_c_fun_demand,carbon_c_fun_inv(tax$mc),current_c)$value+
  integrate(carbon_a_fun_demand,carbon_a_fun_inv(tax$mc),100)$value+
  integrate(carbon_b_fun_demand,carbon_b_fun_inv(tax$mc),100)$value

# (3) No tax revenue
```

Based off of these prices, Sectors A and B will reach equilibriums at `r a_cnt_demand` and `r b_cnt_demand` tons, respectively. Sector C however has a demand of `r c_cnt_demand` tons, therefore they will purchase the remaining credits from Sectors A and B. 

4. Again, without any co-benefits, suppose that country Y (which only has one carbon-emitting sector, D) has no obligation to reduce its emissions. Country X asks country Y to enter the country X carbon market. Doing so would require country Y to put a cap on carbon emissions at its current level of emissions (300 tons), but would then allow country Y to sell carbon offsets to sectors A, B, and/or C. Are there any incentives for country Y to enter country X’s carbon market and to thus place a voluntary cap on its emissions? Are there any incentives for country X to try to attract country Y into its market? Describe these incentives as quantitatively as possible.

```{r}
# Carbon_agg as calculated above, but with sector d included
carbon_agg_d <- data.frame(mc= mc_range,
  sector_a = ifelse(carbon_a_fun_inv(mc_range)>0,carbon_a_fun_inv(mc_range),0),
  sector_b = ifelse(carbon_b_fun_inv(mc_range)>0,carbon_b_fun_inv(mc_range),0),
  sector_c = carbon_c_fun_inv(mc_range),
  sector_d = ifelse(carbon_d_fun_inv(mc_range)>0,carbon_d_fun_inv(mc_range),0)) %>% 
  mutate(agg = sector_a+sector_b+sector_c+sector_d)


# named tax_d for convention but really gives the socailly optimal 
# price of carbon under this scenario. There is a decrease in the 
# price meaning higher reductions possibly?
tax_d <- carbon_agg_d %>% filter(round(agg)==600) %>% 
  summarize(mc=mean(mc),
            sector_a=mean(sector_a),
            sector_b=mean(sector_b),
            sector_c=mean(sector_c),
            sector_d=mean(sector_d),
            agg=mean(agg)
            )


# Kazia goes rogue
# here goes
# kazia has thought about this but gotten nothing completed because # she must go to dinner
# happy to chat through this after dinner if that's helpful Keene

# Below here, Keene pulled in stuff from the last section and hoped that it was the right tactic.  I think this means that credits for 122 tons of emissions would be sold.

#Demand for credits
d_cnt_demand <- carbon_d_fun_inv(tax_d$mc)

#Surplus or deficit of credits
d_cnt_sold <- 300 - d_cnt_demand

# (1) Total Cost
# This is similar to the tax question
# total_cost_cnt <-integrate(carbon_a_fun_demand, carbon_a_fun_inv(tax$mc),current_a)$value+
#   integrate(carbon_b_fun_demand,carbon_b_fun_inv(tax$mc),current_b)$value+
#   integrate(carbon_c_fun_demand,carbon_c_fun_inv(tax$mc),current_c)$value

# (2) Sector cost
# This is similar to the tax question. Add the extra area to calculation such that the sold/purchased credits are accounted for
sector_d_cost_cnt <-  integrate(carbon_d_fun_demand,carbon_d_fun_inv(tax_d$mc),current_d)$value-
  integrate(carbon_d_fun_demand,carbon_d_fun_inv(tax_d$mc),300)$value
```

